---
title: "The browsR"
author: "Tobias Straub"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Genome coverage plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, warning=F, message=F}
library(tsTools)
require(TxDb.Dmelanogaster.UCSC.dm6.ensGene)
library(RColorBrewer)
library(GEOquery)
library(rtracklayer)
```

##Installation

`tsTools` is available on github (https://github.com/musikutiv/tsTools). Documentation is provided as Vignette.

```{r, eval=F}
install.packages("devtools")
library(devtools)
install_github("musikutiv/tsTools", build_vignettes=T)

library(tsTools)
browseVignettes("tsTools”)
```



the function `plotProfiles` takes the following parameters (bold = required)

| name | explanation |
|------------|---------------------------------------------------------------------------------|
**fstart** | start of the genomic window (bp)
**fend** | end of the genomic window (bp)
**fchr** | chromosome of the genomic window
**profs** | named list of coverages i.e. list of SimpleRleLists one per profile to be plotted
cols | colors of the profiles (= vector of length profs)
ann | a named list dataframe of annotations. Dataframes with colnames chr, start, end, col. The col column specifies color if individual data points
ylabel | the label of the y axis
ylims | list of ylimits for plotting the coverages. list(c(min, max), c(min, max))
txdb | a transcription database used for plotting gene models (TxDb format)
ftitle | title of the plot
collapse | collapse gene models (default = TRUE)
with.genes.highlited | vector of gene ids that should be highlighted
plot.labels | plot labels for genes (default = TRUE)
grid | plot grid (default = FALSE)
with.average | plot average (default = FALSE)

## a simple example, one profile

I load the profile from the packages. For sake of space the profile just consists of one chromosome. Therefore I have to explictly re-create a SimpleRleList object. Typically a Granges to coverage conversion would yield a SimpleRleList for direct usage.

The genomic annotations are taken from Bioconductors `TxDb.Dmelanogaster.UCSC.dm6.ensGene` library. Alternatively it can be created from GTF or GFF files using the `GenomicFeatures` package.

```{r, message=F, fig.width=6, fig.height=2}

cov <- new("SimpleRleList")
fpath <- system.file("extdata", "covx.rds", package="tsTools")
cov[["chrX"]] <- readRDS(fpath)

plotProfiles(fstart=1660000,fend=1720000,fchr="chrX", profs=list(MSL2=cov), cols=c("red"),txdb=TxDb.Dmelanogaster.UCSC.dm6.ensGene)

```

## three profiles

In this example I simply re-use the same profile

```{r, fig.height=3, fig.width=6}

cov <- new("SimpleRleList")
fpath <- system.file("extdata", "covx.rds", package="tsTools")
cov[["chrX"]] <- readRDS(fpath)

plotProfiles(fstart=1660000,fend=1720000,fchr="chrX", profs=list(MSL2=cov, MSL3=cov, MSL4=cov), cols=brewer.pal(3, "Set1"),txdb=TxDb.Dmelanogaster.UCSC.dm6.ensGene)

```

## adding annotations

```{r, fig.height=2, fig.width=6}
fpath <- system.file("extdata", "msl2.bed", package="tsTools")
peaks <- read.delim(fpath, header=F)
peaks[,1] <- paste0("chr", peaks[,1])

ann <- list("MSL2 peaks"=data.frame(chr=peaks[,1], start=peaks[,2], end=peaks[,3], col=c("blue")))

plotProfiles(fstart=1660000,fend=1720000,fchr="chrX", profs=list(MSL2=cov), cols=c("red"),txdb=TxDb.Dmelanogaster.UCSC.dm6.ensGene, ann=ann)
```

## getting profiles from GEO (GSM1941084)

Using the `geoQuery` package download a bedgraph file from GEO. `rtracklayer` is then used to import the track which yieals a `GRanges` list. Which is then converted to a `SimpleRleList` required for the `plotProfiles` function. The chromosome names of the coverage has to be adjusted as they don't match the ones of the txdb annotation. 

```{r, fig.height=2, fig.width=6, message=F, warning=F}
filePaths = getGEOSuppFiles("GSM1941084")
granges <- import(rownames(filePaths)[1])
granges
cov <- coverage(granges, weight = "score")
names(cov) <- paste0("chr",names(cov))
plotProfiles(fstart=1660000,fend=1720000,fchr="chrX", profs=list(MSL2=cov), cols=c("red"),txdb=TxDb.Dmelanogaster.UCSC.dm6.ensGene, ann=ann)
```


## Alternatives

https://github.com/rstats-gsoc/gsoc2017/wiki/Interactive-Genome-Browser-in-R
http://www.bioconductor.org/packages/release/bioc/html/GenomeGraphs.html
https://bioconductor.org/packages/release/bioc/vignettes/Gviz/inst/doc/Gviz.pdf
https://bioconductor.org/packages/release/bioc/html/ggbio.html

```{r}
sessionInfo()
```

